---
title: "Module 4 challenge"
author: "Mel Zarate"
date: "9/13/2019"
output: html_document
---

#*Module 4*

#Install package
```{r}
library(stringr)
```

## Matrix and Array practice

```{r}
m <- matrix (data=c(1,2,3,4), nrow=2,ncol=2)
m
```
Using byrow=FALSE will make it go in order of columns
```{r}
m <- matrix(data=c(1,2,3,4,5,6),nrow=2,ncol=3,byrow=FALSE)
m

n <- matrix(data=c(1,2,3,4,5,6),nrow=2,ncol=3,byrow=TRUE)
n
```
Binding vectors using rbind() function (by rows) or cbind() (by column)
```{r}
v1 <- c(1,2,3,4)
v2 <- c(6,7,8,9)
m1 <- rbind(v1,v2)
m1
m2 <- cbind(v1,v2)
m2
```
Extracting metadat from matrices
```{r}
class(m1)
dim(m1)
colnames(m2)
str(m2)
```
Array: can be 3 dimenisonal
dim=c(5,6,3) means it has 5 rows, 6 columns and 3 levels
```{r}
a <- array(data=1:90, dim=c(5,6,3))
a
```
## Subsetting 

Use brackets to select elements from vector, matrices or arrays
```{r}
v <- 1:100
v
v[1:15]
```
specific values: 
```{r}
v[c(2,4,6,8,10)]
```
Subset by frunction 
```{r}
v[seq(from = 1, to = 100, by = 2)]
```

###CHALLENGE###

Creating a vector and extracting every third word 
```{r}
v <- c("Stately", "plump", "Buck", "Mulligan", "came", "from", "the", "stairhead", "bearing", "a", "bowl", "of", "lather", "on", "which", "a", "mirror", "and", "a", "razor", "lay", "crossed")
v
v[seq(1, length(v), 3)]
```
##Challenge 1##
I did not know the difference between the inclusion and exclusion of quotations. When I first tried creating the vector without quotations, any type of subset was coming back as the entire statement. 
using stringr
```{r}
str_split(v, " ") #splits all elements
```
##Challenge 2##
Looking online, I figured out how to split the vector starting and ending at different elements or by different characters, but could not manage to do it by nth character. 

Creating and extracting elements from matrices
```{r}
m <- matrix(data=1:80,nrow=8,ncol=10,byrow=FALSE)
m
x <- m[4,5]
x
```
### CHALLENGE 2
extract columns
```{r}
x <- m[ ,c(2,3,6)]
x
```
extract rows
```{r}
x <- m[6:8,]
x
```

```{r}
x <- m[c(2, 2), c(6, 9)]
x #this didn't work, only gave overlapping elements
x <- m[10:70]
x
```
##Challenge 3## 
Could not find the input to subset from one specific element to another; I was only getting subsets of elements that share certain rows and columns.

###CHALLENGE 3
```{r}
a <- array(data=400:1, dim=c(5,5,4,4))
a
```
```{r}
a[1,1,1,2]
a[2, 3, 2, ]
a[1:5, 1:5, 3, 3]
```
Replacing elements in matrix m 
```{r}
m[7,1] <- 564
m[,8] <- 2
m[2:5,4:8] <- 1
m[2:5,4:8] <- c(20,19,18,17)
m[2:5,4:8] <- matrix(data = c(20:1), nrow=4,ncol=5,byrow=TRUE)
m[,8] <- c("a","b")
m
```
making a list 
```{r}
s <- c("this","is","a","vector","of","strings")
m <- matrix(data=1:40, nrow=5, ncol=8)
b <- FALSE
l <- list(s, m, b) 
l
# extracting elements from the list
l[[2]] #double brackets to choose a whole element (data type- here we are choosing matrix)
l[[2]][2,6] #more specific- row 2, col 6 of matirx
l[1:2] #looking at string and matrix
l[c(1,3)]
str(l) #gives details about list 
names(l) <- c("string", "matrix", "logical") #give names to each item
names(l)
l$logical #use name to see it 
l$matrix[3,5] #row 3 col 5 in the matrix
```
### CHALLENGE 4
Work up from lowest taxonomic level...
```{r}
#first the haps!
Catarrhini<-list("Cercopithecidae", "Hylobatidae", "Hominidae")
names(Catarrhini)<-c("Cercopithecidae", "Hylobatidae", "Hominidae")
Platyrrhini<-list("Cebidae", "Atelidae", "Pitheciidae")
names(Platyrrhini)<-c("Cebidae", "Atelidae", "Pitheciidae")
Anthropoidea<-list(Catarrhini,Platyrrhini)
Tarsioidea<-"Tarsiidae"
Haplorhini<-list(Anthropoidea,Tarsioidea)
Haplorhini
#streps!
Lemuroidea<-list("Cheirogaleidae","Lepilemuridae","Indriidae","Lemuridae","Daubentoniidae")
names(Lemuroidea)<-c("Cheirogaleidae","Lepilemuridae","Indriidae","Lemuridae","Daubentoniidae")
Lorisoidea<-list("Lorisidae","Galagidae")
names(Lorisoidea)<-c("Lorisidae","Galagidae")
Strepsirhini<-list(Lemuroidea,Lorisoidea)
names(Strepsirhini)<-c("Lemuroidea","Lorisoidea")
Strepsirhini
#put them together!
Primates<-list(Strepsirhini,Haplorhini)
Primates
```
#data frame
Create from scratch:
```{r}
df <- data.frame(firstName=c("Nitin","Silvy","Ingrid","Claire"), program=c("iSchool","Anthro","Anthro","EEB"), sex=c("M","F","F","F"), yearInProgram=c(1,2,2,3))
df
```

read from user's data:
```{r}
df <- read.table("~/Desktop/random_people.txt", sep =",", header = TRUE, stringsAsFactors = FALSE)
df[,c(1,3,4,11,12)]
str(df)
df[,3] #subset data 
str(df[,3])
```
```{r}
df$name.last
str(df$name.last)
```

```{r}
df[3]
str(df[3])
```
```{r}
df["name.last"]
str(df["name.last"])
```
adding rows(cases)/columns(variables)
```{r}
df <- cbind(df, id=c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20)) #cbind adds a column (with 20 rows?)
df <- cbind(df, id=c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20)) #giving identities to each row
df
```
doing so using $
```{r}
df$school <- c("BU","BU","Harvard","Harvard","BU","Boston College","MIT","BU","BU","UMASS Boston","Harvard","BU","Boston College","BU","Harvard","MIT","Harvard","BU", "UMASS Boston", "Harvard")
df
```
Using a logical vector to subset:
```{r}
new_df <- df[df$school == "BU",]
new_df
```
returns a logical vector equal in length to the number of rows in the original data frame, then subsets df based on that vector giving only values TRUE (matching parameters)

Return only particular columns when we subset:
```{r}
new_df <- df[df$school == "BU", c("name.last", "name.first", "school")]
new_df
```

###CHALLENGE 5
```{r}
m <- matrix(data=c(3,0,1,23,1,2,33,1,1,42,0,1,41,0,2), nrow=5,ncol=3,byrow=TRUE)
m
```
coerce into a data frame:
```{r}
df <- as.data.frame(m)
df
```
Coerce second column to be logical-valued
```{r}
as.logical(df$V2)
```
Coerce third column to be factor-valued
```{r}
as.factor(df$V3)
```
show data type for each variable
```{r}
str(df)
```
##Challenge 4## 
I'm not quite sure what exactly the str() function is showing me and if it alingns with the changes I made to the data frame. 

##Challenge 5##
Pushing to GitHub: I had a few challeges here. First, I had trouble getting the markdown that I was working on to show up in the "git" tab in R studio. I eventually re-opened the project and made a new Rmd that I then copied my work from my previous markdown to. The obvious was to avoid this next time is to open the markdown within that project initially. I am also experiencing problems pushing the knitted html (or anything else) to GitHub, as it is giving me an error message pertaining to access to my URL everytime.